\documentclass[a4paper]{exam}

\usepackage[a4paper]{geometry}
\usepackage{hyperref}
\usepackage{xcolor}

\header{CS/MATH 113}{WC06: Logical Inference}{Spring 2024}
\footer{}{Page \thepage\ of \numpages}{}
\runningheadrule
\runningfootrule

\printanswers

\newcommand\cb{\color{blue}}
\newcommand\cre{\color{red}}
\newcommand\cg{\color{green}}
\newcommand\cy{\color{yellow}}
\newcommand\cgr{\color{gray}}

\title{Weekly Challenge 06: Logical Inference}
\author{CS/MATH 113 Discrete Mathematics}
\date{Spring 2024}

\qformat{{\large\bf \thequestion. \thequestiontitle}\hfill}
\boxedpoints

\begin{document}
\maketitle

\begin{questions}

  \titledquestion{Sacred Secrets}[5] Legend has it that one of the TAs from a previous offering of CS/MATH 113 prepared a manual, ``Sacred Secrets: How to Earn an A+ and Stay Sane''. However their own sanity is sadly no longer intact. The \LaTeX\ source and the repository got deleted and all that exists about the location of the only printed copy are the following instructions.
  \begin{enumerate}
  \item There is a hint at Learn Courtyard or at the Gym.
  \item If your TA is sitting in Ehsas or they are absent, then there is a hint at Learn Courtyard.
  \item If your TA is not sitting in Ehsaas, then there is a hint at the Gym.
  \item If there are people in Learn Courtyard, then there is no hint at Learn Courtyard.
  \item If there is a hint at Learn Courtyard, then the manual is at Zen Garden.
  \item If there is hint at the Gym, then the manual is at Earth Courtyard.
  \item If your TA is absent, then the manual is at Fire Courtyard.
  \end{enumerate}
  You notice that there are people in Learn Courtyard. Show how you can infer the location of the manual.

  \begin{solution}\\\\
    % Enter your solution here.
    \begin{tabular}{|cc|c|}
      \hline
      & step & reason \\ 
      \hline
      8) & There are people in Learn Courtyard & \\
      9) & There is no hint at Learn Courtyard & Modus ponens (4)(8) \\
      10) & TA is not sitting in ehsas and they are not absent & Modus tollens (2)(9)\\
      11) & TA is not sitting in ehsas & simplification (10)\\
      12) & There is a hint at the gym & Modus ponens (11)(3)\\
      13) & The manual is at Earth Courtyard & Modus ponens (12)(6)\\
      \hline
    \end{tabular}\\
    \\Therefore the location of the manual is Earth courtyard found by a series of inferences
  \end{solution}

\titledquestion{The Other Side} You are given four cards each of which has a number on one side and a letter on another. You place them on a table in front of you and the four cards read: \texttt{A 5 2 J}. Which cards would you turn over in order to test the following rule? Explain your choice.
      {\quotation Cards with 5 on one side have J on the other side.}

  \begin{solution}
    % Enter your solution here
    Since we are given that 5 on one side should have J on the other, it is possible that there is J with another number but for every card with 5 there should be J on the other.\\
    The card turned over to check this rule should be 5.\\

    V(x): The character on visible side of card x is 5 \\
    C(y): The character on the side turned over of card x is J\\
    The rule can be universally quantified as: $\forall x (V(x) \rightarrow C(x))$
    
  \end{solution}

  
\end{questions}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
